---
title: "Homework5"
author: "LeTicia Cancel"
date: "9/25/2020"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message=FALSE, warning=FALSE}
#Libraries needed
library(RCurl)
library(dplyr)
library(tidyverse)
library(tidyfast)
library(tidyr)
library(ggplot2)
```

## Import Flights Data
Import Flights data from GitHub
```{r}
site <- getURL("https://raw.githubusercontent.com/ltcancel/Homework-5/master/Airline%20Data.csv")
df <- read.csv(text = site, sep = ",", header  = TRUE)

head(df)
```

## Tidy the data
Some of the columns need to be renamed. The data also imported with null values where a row will need to be removed or observations will need to be filled with the appropriate data. 
```{r}
#Fix column names of the first two columns
df <- rename(df, Airline = X, Status = X.1)
df

#Remove the null row
df <- df %>% drop_na()
df
```
I tried to use the fill function to fill in the Airline data in the observations that are blank but it did not work at first. R did not read these observations as NULL or NA so fill did not work. I made these observations NA so fill would work. Even though I could have just filled in the observations with the exact values, I chose to go this route to practice using some of these functions. 
```{r}
#Make blank Airline NULL so it can be filled
df[2,1] <- NA
df[4,1] <- NA

df


#Fill the new NA columns with the values from the row above
df <- fill(df, Airline)
df


#Create a new data frame making some of the columns as rows instead
long_df <- df %>% dt_pivot_longer(
  cols = -c(Airline,Status),
  names_to = "City",
  values_to = "Flights"
)

head(long_df)
```



## Flights Analysis

Analysis to compare arrival delays for both airlines. 
```{r}
#split data by flight status 
onTime <- long_df %>%
  filter(Status == "on time")

delayed <- long_df %>% 
  filter(Status == "delayed")

onTime
delayed

#combine subsets so each flight status is in their own column
final_flights <- cbind(onTime, delayed[,4])
final_flights

#Tidy final data frame to keep only the columns needed and rename columns
final_flights <- final_flights %>%
  rename(
    onTime = Flights,
    Delayed = `delayed[, 4]`
  ) %>%
  select(Airline,City,onTime,Delayed)

final_flights

#Comparison of on time flights by Airline for each city
ggplot(final_flights, aes(x = Airline, y = onTime, fill = Airline)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~City)

#Comparison of delayed flights by Airline for each city
ggplot(final_flights, aes(x = Airline, y = Delayed, fill = Airline)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~City)


#Comparison of cumulative status for each airline
ggplot(long_df, aes(x = Status, y = Flights, fill = Airline)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~Airline)
```

